---
title: "Homework 1 -- Version 2"
author: "YOUR NAME""
date: "09/26/2014"  
output: html_document
---
To prepare for this homework, do the following: 

1. On line 3 of this document, replace "YOUR NAME" with your name. 
2. Place your cursor on line 12 of this document and press CTRL-Enter:
```{r}
source('http://www.stat.cmu.edu/~nmv/setup/stat202/hw-1.R')
``
3. Look at the output from the console. 
    a. If the outputs ends with "Done.", then proceed to Step 4. 
    b. If the output asks you to install a package, do so, then return to Step 2. 
    c. If you get a different error, email the text of error to 
       nathanvan AT northwestern FULL STOP edu. 
4. Click the **Knit HTML** button. Be patient, it may take a minute to complete. 
5. If the 'knitting' worked, you will see a pretty version of this document pop up.
6. Close the pretty version and come back to the plain text to do the assignment.
7. After every question, click the **Knit HTML** button to make sure that the output is as expected. 
8. After finishing the assignment, hand in the knitted 'stat-202-hw-1-version-2.html' file via [BlackBoard](https://courses.northwestern.edu/).

## Introduction 

The 'stat-202-hw-1-version-2.Rmd' file that you are currently reading is the **source file** 
that RStudio uses to produces the pretty 'stat-202-hw-1-version-2.html' file. 

An '.Rmd' file combines three types of information into one place: 

1. Prose. 
2. Formatting information so that some words are **bold**, others are *italic*, etc. Go [here](http://rmarkdown.rstudio.com) for further details. 
3. R code and graphs. 

R code inside of '.Rmd' files is encapsulated in R code "chunks". For example, the code in Step 2, on lines 11 to 13, was an example of an R code chunk. When you clicked the **Knit HTML** button, RStudio (i) applied the formatting information to the prose, (ii) ran the R code in the chunk, and then (iii) 'knit' together the formatted text, R code, and R output into the pretty 'stat-202-hw-1-version-2.html' file that was then immediately opened in RStudio's browser. 

Note that the output displayed in the '.html' file is the same as the output that appeared on the console when you run it with CTRL-Enter. This is what makes '.Rmd' files powerful: you can work on the homework by executing commands with CTRL-Enter, and after you are satisfied with the output you can knit it together to turn it in. 

## A short intro to R
In this section we will practice using R inside of an '.Rmd' file. Follow the 
instructions in the following code chunk. (Note: the R code will not appear in the knitted output because of passing `echo=FALSE` and `eval=FALSE` in the setup of the code chunk. Do not be concerned.)

```{r, echo=FALSE, eval=FALSE}
# In an R code chunk, the pound sign indicates that the line you are reading
# is a comment. You can tell this at a glance because RStudio automatically colors
# comments green. You should read all of the comments in the file. 

# R can be used as a simple calculator. 

# Put your cursor on the next line. Hit CTRL-Enter.
1+1

# In the console below this screen you should see: 
# > 1+1
# [1] 2
# >

# Try something that looks a bit more like algebra. Again, put your 
# cursor on the next line and hit CTRL-Enter. 
x <- 1+3

# This code means: add 1 plus 3 and store the result in x. 
# What you should see is: 
# > x <- 1+3
# > 
# Note that in this case, R doesn't output anything on the console. No output
# usually means things worked as expected. 

# So how do we access the value of x? 
# Try hitting CTRL-Enter on the next line.
x 

# This gets you: 
# > x 
# [1] 4

# Now R implements the standard PEMDAS order of operations. Note that
# the two expressions are not equivalent to each other. (You know what 
# to do from now on with CTRL-Enter.)
x^2/4
x^(2/4)

# R can also work with vectors and matrices.
# Let's generate a vector that counts from 1 to 9. 
1:9

# That gives you
# > 1:9
# [1]  1  2  3  4  5  6  7  8  9

# Let's store that in a variable:
vectorA <- 1:9

# Predict what the following expressions will do: 
vectorA + 10
3 * vectorA
vectorA^2

# Did you get them right? 
# Answers: 
# ... Add 10 to each element in the vector
# ... Multiply each element in the vector by 3
# ... Square each element in the vector 

# Of prime importance for this course are the statistical functions 
# that we can use. For example, what do these do? 
mean(vectorA)
median(vectorA)

# Answers:
# ... caculates the mean
# ... caculates the median
```

## Assignment 
As discussed in class the Behavioral Risk Factor Surveillance System (BRFSS) is an annual telephone survey of people living the United states. Further information is available [from the CDC](http://www.cdc.gov/brfss/).

A 20,000 record subset of the 2013 BRFSS is provided by [OpenIntro Statistics](http://www.openintro.org/) for public use. The setup script that you ran in Step 2 loaded that subset into the variable `cdc`. 

Use the R code chunk below to explore the data and make a few charts. As before, this code will not be present in the knitted output. 

```{r, echo=FALSE, eval=FALSE}
# View the first six rows of the data 
head( cdc )

# View the last six rows of the data
tail( cdc )

# Explore the data in a spreadsheet.
#  *** Warning **** 
#   Running the following command will HIDE these instructions. 
#   Click on the 'stat-202-hw-1-version-2.Rmd' tab to return to these 
#   instructions after you explored the spreadsheet.
View( cdc )

# For the genhlth column ... 
#
# ... make a table of raw counts 
table( cdc$genhlth )
#
# ... make a table of proportions 
# ... Step 1: How many records? 
nrow(cdc)
# ... Step 2: Divide the raw counts by the number of records
table( cdc$genhlth ) / nrow( cdc )
#
# ... make a pie chart 
#     *** Warning *** 
#     The pie chart appears in the bottom left window of RStudio.
#     If it is too small to see well, click the "Zoom" button for that 
#     window. 
pie( table(cdc$genhlth) )
#
# ... NB. The pie function *requires* a table as input
#         If you attempt to run 
#             pie( cdc$genhlth )
#         you will get a cryptic error message. 
#
# ... make a bar plot 
barplot( table( cdc$genhlth ) )
#
# ... and put the pie chart and the barplot side by side
par(mfrow=c(1,2))
  pie( table(cdc$genhlth) )
  barplot( table( cdc$genhlth ) )
par(mfrow=c(1,1))

# And for the height column ... 
# 
# ... make a histogram 
hist( cdc$height  )
# 
# ... add more bars to the histogram by increasing the 
#     number of breakpoints 
hist( cdc$height, breaks=25 )
#
# ... find the mean
mean( cdc$height )
#
# ... find the median 
median( cdc$height )
#
# ... find the five number summary
summary( cdc$height )
#
# ... making a boxplot
boxplot( cdc$height )
```


### Question #1 

The columns of the `cdc` data frame are:

 * `genhlth`  What is your general health? 
 * `exerany`  Do you ever exercise? (0 = No, 1 = Yes)
 * `hlthplan` Do you have insurance? (0 = No, 1 = Yes)
 * `smoke100` Have you smoked 100 cigarettes in your life? (0 = No, 1 = Yes)
 * `height`   How tall are you in inches? 
 * `weight`   How much do you weigh in pounds? 
 * `wtdesire` How much would you like to weight? 
 * `age`      What is your age in years? 
 * `gender`   Are you male or female? 

WHICH VARIABLES are quantitative and which are categorical? 

### Question #2

Make a pie chart and a bar plot for gender:
```{r}
par(mfrow=c(1,2)) # <- leave this line

  ## Code for the pie chart: 
  pie( table( cdc$gender)  )

  ## Code for the bar chart goes below here: 
  barplot( table(cdc$gender ))

par(mfrow=c(1,1)) # <- leave this line
```

REPLACE this message with a short description of the distribution of gender in the sample. 

COMMENT on which display you prefer and **why**. You are allowed to say pie chart, it is not a trick question. Always support your opinion with an explanation. 

### Question #3 

Choose a categorical variable besides `genhlth` and `gender` in the `cdc` data set. Make a pie chart and a bar chart for that variable: 
```{r}
par(mfrow=c(1,2)) # <- leave this line

  ## Code for the pie chart: 

  ## Code for the bar chart goes below here: 

par(mfrow=c(1,1)) # <- leave this line
```

REPLACE this message with a short description of the distribution of the variable. 

COMMENT on which display you prefer and **why**. 

### Question #4

Make a histogram for `height`:  
```{r}
# Code for the histogram goes below this line: 


``` 

DESCRIBE the distribution of height in this sample (e.g. shape, center, spread, and deviations from the pattern). 

Calculate the mean and the median of `height`. 
```{r}
# Code to calculate the mean goes below this line: 

# Code to calculate the median goes below this line: 


```

EXPLAIN why the values of the mean and median are either (i) the same or (ii) different. 

WHICH is better measure of center for these data?

### Question #5

Make a histogram and boxplot for `wtdesire`:  
```{r}
par(mfrow=c(1,2)) # <- leave this line

# Code for the histogram goes below this line: 

# Code for the boxplot goes below this line: 


par(mfrow=c(1,1)) # <- leave this line
``` 

COMPARE AND CONTRAST the two displays. 

DESCRIBE the distribution of desired weight in this sample (e.g. shape, center, spread, and deviations from the pattern). 

Calculate the mean and the median of `wtdesire`. 
```{r}
# Code to calculate the mean goes below this line: 


# Code to calculate the median goes below this line: 



```

EXPLAIN why the values of the mean and median are either (i) the same or (ii) different. 

WHICH is better measure of center for these data?

### Question #6
Generate a five number summary for `height`. 

```{r}
# Code to calculate the five number summary goes below this line: 


```

The following code generates a side-by-side boxplot for each gender's height. 
```{r}
# Read the ~ sign as "by", i.e. a boxplot of height by gender of the cdc data: 
boxplot(height ~ gender, data=cdc)

```

WHAT is revealed by the side-by-side boxplot that is hidden in the five number summary? 


### Question #7
In this question we will compare a person's reported health status with how much weight a person wants to lose. 

```{r}
## Step 1: Add a column to the cdc data frame for desired weight loss 
cdc$wtloss <- cdc$wtdesire - cdc$weight 

## Step 2: Generate boxplots of desired weight loss by general health status 
## ... Note: the ylim argument zooms in the boxplot to the range -100 to 100 lbs.
boxplot( wtloss ~ genhlth, data = cdc, ylim=c(-100,100))

## Step 3: Add a horizontal line at zero weight loss desired 
abline( h = 0 )

```

WRITE A PARAGRAPH (at least 5 sentences!) INTERPRETING this graph. 
