\documentclass[10pt]{article}

\usepackage{outlines}
\usepackage[margin=1in]{geometry}
\usepackage[hidelinks]{hyperref}


\title{Statistics 202: Introduction to Statistics}
\author{Spring 2014}
\date{\today}


\begin{document}
\maketitle

\section*{Instructors}
Nathan VanHoudnos (van-HOD-ness), Ph.D. \\
Institute for Policy Research, 2046 Sheridan Road, Room 304A. \\
{\bfseries Office Hour: Tuesday 10-11 AM in my office}  \\
Email: \href{mailto:nathanvan@northwestern.edu}{\nolinkurl{nathanvan@northwestern.edu} }\\
Info: \url{http://www.stat.cmu.edu/~nmv} \\

\noindent TA: Aaron Kleyn \\
Email: aaronKleyn2018@u.northwestern.edu \\
Office Hours: Thursday 10-11 AM in the stats lounge 

\section*{Course information}

\begin{outline}

\1 {\bfseries Description:} Via the course catalog:
\begin{quote}
{\bfseries 202 Introduction to Statistics }

Data collection, summarization, correlation, regression, probability,
sampling, estimation, tests of significance. Does not require calculus
and makes minimal use of mathematics.
\end{quote}

\1 {\bfseries Objective:} Via the syllabi of past instructors: 
\begin{quote}
To help you develop a critical attitude toward statistical
arguments. This course is for people who want to be able to comprehend
and use statistics better in their work. This course stands by itself
and also serves as a background for further statistics courses,
helping to provide the intuition which can sometimes be lost amid the
formulas. 
\end{quote}

\1 {\bfseries Prerequisites:} High school algebra

\1 {\bfseries Website:} \url{https://courses.northwestern.edu}

\1 {\bfseries Required Textbook:} {\em Probability and Statistics} by
the Open Learning Initiative (OLI) of Carnegie Mellon University. 

   \2 The OLI textbook is online only. It is accessed via our course
   BlackBoard site ( \url{https://courses.northwestern.edu} ) under
   the Assignments section. 

   \2 The textbook is purchased directly from OLI upon first access
   (\$25). 

\1 {\bfseries Course Format:} This course uses a hybrid of traditional
lecture based and online instructional strategies. 

   \2 {\bfseries Lecture:} MWF from 10-10:50 AM in Annenberg Hall
   G15. Lecture serves to (i) introduce new material, and (ii) address
   common student questions. Attendance is mandatory. (3 hours / week) 

   \2 {\bfseries Reading and ``Did I get this?'':} The OLI readings
   expand on the material presented in lecture and offer
   opportunities for checking your understanding with {\em ungraded}
   ``Did I get this?'' activities. ( approx. 4 hours / week ) 

   \2 {\bfseries ``Checkpoint'' quizzes:} These OLI quizzes cover
   material from the previous lecture and are due at 9 AM before the
   next lecture. ``Checkpoint'' quizzes are graded
   pass/fail. {\bfseries Quizzes are to be soley your own work.} (approx. 45 minutes / week )

   \2 {\bfseries Homework:} Homework assignments are due
   on Tuesdays at noon. (approx. 3 hours / week )

       \3 Homework must be handed in via BlackBoard in the specified format. Non-conforming homework will not be accepted. 

       \3 Your name, and the name of {\em everyone} who contributed to
       your solutions, must appear on your homework. 

       \3 Late homework will not be accepted.

       \3 The lowest scoring homework assignment will be dropped
       before calculating grades.
  
   \2 {\bfseries Exams:} According to Weinberg College of Arts and
   Sciences policy, every student must take the final exam at its
   scheduled time. All midterm conflicts must be resolved in advance.
       
       \3 Midterm: October 24, 2014 in class 
       \3 Final: December 10, 2014 from 3-5 PM

\1 {\bfseries Grading:}
  \2 15\% Checkpoints 
  \2 25\% Homework
  \2 30\% Midterm  
  \2 30\% Comprehensive final 

\1 {\bfseries Laptop policy:} A computer is required for the course,
but do not bring your laptops to lecture unless requested. If you
require or strongly prefer a laptop for note taking, please see me. 

\1 {\bfseries Calculator policy:} A calculator is required for
exams. Graphing calculators are not necessary. The use of phones and similar
devices is forbidden during exams. Calculators may not be shared
during exams.  

\1 {\bfseries AccessibleNU:} Any student requesting accommodations
related to a disability or other condition is required to register
with AccessibleNU (accessiblenu@northwestern.edu or 847-467-5530) and
provide professors with an accommodation notification from
AccessibleNU, preferably within the first two weeks of class. All
information will remain confidential.

\1 {\bfseries Academic Integrity:} Suspected violations of academic
integrity will be reported to the Dean's Office. For more information
on Northwestern’s academic integrity policies, see
\url{http://www.weinberg.northwestern.edu/handbook/integrity/index.html}.

\end{outline}

\break
\section*{Stat 202 Schedule} 
Lecture A-B denotes that lecture will cover pages A-B in the
textbook. Checkpoints are due at 9 AM on the indicated day. Homework
is due at noon on the indicated day. Class sessions marked
``Instructor Review'' and ``Student Review'' are designated in-class
review sessions that are led by the instructor and students
respectively. Class sessions marked ``Special Topics'' will cover
additional topics of interest to the students as time allows. In
accordance with Weinberg College of Arts and Sciences policy, no
assignments are due during reading week and attendance at class
meetings during that week is strictly optional. \\

\begin{tabular}{c|c|c|c|c|c}
\hline
Week beginning & Monday & Tuesday & Wednesday & Thurs. & Friday \\
\hline
9/22  & $\times$ & $\times$ & Lecture 1-21 & $\times$ &  Checkpoint 1 \\
      &          &          &              &  &  Lecture 22-33       \\
\hline 
9/29  & Checkpoint 2   & Homework 1  & Checkpoint 3   & $\times$ &  Checkpoint 4 \\
      & Lecture 34-47  &             & Lecture 48-61  &  &  Lecture 62-78   \\
\hline 
10/6  & Checkpoints 5 \& 6  & Homework 2  & Checkpoint 8   & $\times$ &   Checkpoint 9 \\
      & Lecture 84-99  &              & Lecture 100-111 &  &   Lecture 112-115  \\
\hline 
10/13  & Checkpoint 10  & Homework 3  & Checkpoint 11   & $\times$ &  Checkpoints
12 \& 13 \\
      & Lecture 116-123 &             & Lecture 124-140 &  &   Lecture 145-155  \\
\hline 
10/20  & Checkpoint 15      & Homework 4  & Student Review  & $\times$ &  {\bfseries Midterm 10/24}  \\
       & Instructor Review  &             &                 &  &
{\bfseries in class }          \\
\hline 

10/27  & Lecture 160-165 & $\times$  & Checkpoint 16 & $\times$ &  Checkpoint 17  \\
       &             &           & Lecture 166-169    &  &  Lecture 171-187    \\
\hline 
11/3   & Checkpoint 18A & Homework 5 & Checkpoint 18B & $\times$ &  Checkpoint 19  \\
       & Lecture 188-191   &            & Lecture 192-198   &  &  Lecture 199-211    \\
\hline 
11/10  & Checkpoint 20 & Homework 6 & Checkpoint 21 & $\times$ & Checkpoints 23 \& 24  \\
       & Lecture 212-220   &            & Lecture 222-234   &  &  Lecture 235-242    \\
\hline 
11/17  & Checkpoint 25 & Homework 7 & Checkpoint 26 & $\times$ &
Checkpoint 28A  \\
       & Lecture 243-249   &            & Lecture 251-256   &  &  Lecture 257-259    \\
\hline 
11/24  & Checkpoint 28B   & Homework 8 & Special Topics & $\times$ &
{\em No Class 11/28} \\
       & Special Topics                &            &                &
&  {\em Thanksgiving Break}   \\
\hline 
%% Add in multicolumn
12/1  & {\em Reading Week}  & $\times$ &  {\em Reading Week} &
$\times$ & {\em Reading Week} \\
      & Instructor Review &        & Student Review  &  & Student Review  \\
\hline
12/8  & $\times$ & $\times$ & {\bfseries Final 12/10} & $\times$ &  $\times$ \\
      &          &          & {\bfseries 3-5 PM} & & \\ 
\hline
\end{tabular} \\

A detailed list of the checkpoints follows. 

\break 
\subsection*{Checkpoints}
Checkpoints are numbered sequentially. Checkpoints marked ``skipped''
assess material that is not covered in the class and are not
graded. Checkpoints marked ``On BlackBoard'' are taken directly on
BlackBoard instead of via the OLI software. Checkpoints marked
``practice'' assess material covered in the course, but they are ungraded. \\

\begin{tabular}{l|r|l|c}
  Unit & No. & Checkpoint Name &  Reading \\
\hline
Exploratory Data Analysis 
 & 1 & Examining Distributions \# 1 & 1-21 \\
 & 2 & Examining Distributions \# 2 & 22-33     \\
 & 3 & Examining Relationships \# 1 & 34-47     \\ 
 & 4 & Examining Relationships \# 2 & 48-61     \\ 
Producing Data 
 & 5 & Sampling           & 62-68 \\
 & 6 & Designing Studies \# 1       & 69-78  \\
 & 7 & Designing Studies \# 2 {\bfseries (skipped) } & 79-83 \\
Probability 
 & 8 & Probability \# 1   & 84-99  \\
 & 9 & Probability \# 2   & 100-111  \\
 & 10 & Conditional Prob. and Indep. \# 1 & 112-115 \\ 
 & 11 & Conditional Prob. and Indep. \# 2 & 116-123  \\ 
 & 12 & Random Variables \# 1 & 124-137  \\
 & 13 & Random Variables \# 2 & 138-140  \\ 
& 14 & Random Variables \# 3 {\bfseries (skipped)} & 141-144 \\
 & 15 & Random Variables \# 4 & 145-155,159,170 \\
%% \hline 
%% Probability 
 & 16 & Sampling Distributions \# 1 & 160-165      \\
 & 17 & Sampling Distributions \# 2 & 166-169      \\
Inference 
 & 18A & Estimation \# 1 {\bfseries (on BlackBoard)} & 171-187 \\
 & 18A & Estimation \# 2 {\bfseries (on BlackBoard)} & 188-190 \\
 & 18C & Estimation {\bfseries (practice)} & 191 \\
 & 19 & Hypothesis Testing (HT) Overview  & 192-198  \\
 & 20 & HT for a Population Proportion  &  199-211 \\
 & 21 & HT for a Population Mean  & 212-220 \\
 & 22 & HT {\bfseries (practice)} & 221 \\
 & 23 & Type I and Type II  & 222-223  \\
 & 24 & Two Independent Samples & 224-234 \\
 & 25 & Matched Pairs & 235-242 \\
 & 26 & ANOVA & 243-249  \\ 
 & 27 & Case C$\rightarrow$Q {\bfseries (practice)} & 250    \\
 & 28A & Case C$\rightarrow$C {\bfseries (on BlackBoard)} &  251-256  \\ 
 & 28B & Case Q$\rightarrow$Q {\bfseries (on BlackBoard)} &  257-258  \\ 
 & 28C & Case C$\rightarrow$C and Q$\rightarrow$Q {\bfseries  (practice) } &  259  \\ 
 & 29 & Inference for Relationships {\bfseries (practice) } & 260  
\end{tabular}


\end{document}
